---
title: "lab-07-simpsons.Rmd"
author: "dalal salman"
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham
```
    
Your answer:
The data is observational as the description states that is based on age, smoking, and mortality, which are all observable events and not produced via experiments.

2.  
```{r observations}
nrow(Whickham)
```

Your answer;

Your answer; There ane 1,314 observations. As we know every row is an observation.

3.  

```{r number_variables}
names(Whickham)
```

Your answer:  
There are 3 variables, "outcome" "smoker", and "age"

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)
```

Your answer: 
Using the unique() function on the 3 variables we could see that "outcome" only takes Alive or Dead value, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age is numerical continous data.
continous One of the best ways to visualise categorícal data is through the use of bar charts.

```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +geom_boxplot()
```


4. 
i expact the health will be worser and may be the person will be died after while,if he keeping smoke

```{r plot_smokeVSincome_1}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)

```
smoker(732) NO..... 31.4 (dead)   >>(68.6) Alive
smoker(582) yes....23.8  (dead)   >>(76.2) Alive

i does not expected this result becaus now the most died people not smoker
6.

```{r age_cat}
Whickham <- Whickham %>% mutate(age_cat = case_when(age <= 44 ~ "18-14" , age > 44 & age <= 64 ~ "45-64" , age > 64 ~ "65+"))
 view(Whickham)
```

7. 

```{r plot_smokeVSincome_3}
ggplot(data =Whickham, aes(x=outcome, y= smoker,color=smoker, fill=age_cat)) + geom_bar(stat="identity") +facet_wrap(~age_cat)
```

what changes > the category of the age it's appear to us and we see the most of dead people not smoker in age (65+) .. and in age (45-64)
and(18-44) the most dead pepole is smoker that is relationship between the smoking and helath not clearly but can say that your helath will be change to worst if you be smoker 

Knit, commit, and push to github.